# Estimation Strategies

```{r packages, message=FALSE, echo = F}
library(tidyverse)
library(sf)
library(raster)
library(Matrix)
library(knitr)
library(kableExtra)
library(ggthemes)
library(cowplot)
library(transformr)
library(gganimate)
set.seed(762)
```

We have implemented two specific estimation strategies:

-   Voronoi tessellation (tower locations OR coverage area centroids as seeds)

-   MLE Poisson

All three of these estimation techniques produce a tile specific estimate. The sum of these estimates always equals the sum of the c.vector, i.e. all registered mobile phones. This characteristic explains the strategy of all of these estimators: Taking the complete number of registered phones in the c.vector and allocating them across all the tiles. There is however a fundamental difference between the two estimators: The second one allows for the realistic setting of overlapping antenna coverage. As shown above there are many tiles that are covered by multiple antennas. Current research suggests that this can result in more accurate estimations.

## Voronoi tesselation

Voronoi tessellation uses seeds to calculate Voronoi regions. For every seed there is one Voronoi region that contains the area that is closest to the particular seed. There are two seed specifications we are interested in:

-   Tower locations as seeds --\> n = 1289

-   Coverage area centroid (per antenna) locations as seeds --\> n = 3841

Below we specify the tower locations as seed points. These are then used to compute the Voronoi regions. Before we do that we need to make sure that the corresponding c.vector is aggregated on the tower level. We then join the Voronoi regions with the tiles to create an object that links tiles within respective Voronoi regions.

### Setup of necessary datasets

For any kind of Voronoi tessellation we need to work with the grid level data (tiles) and the cell coverage data (tower, antennas).

```{r data load in, eval=F}
census.de.100m.tile <- readRDS("Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/working objects/census.tile.final.rds") 

coverage.areas <- readRDS("Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/working objects/coverage.areas.rds") 

c.vec <- readRDS("Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/working objects/C.vec.df.final.new.rds") 

## derive the shape of the focus area
# census.geo.body <- census.de.100m.tile %>% 
#   select(internal.id) %>% 
#   summarise(geometry = st_union(geometry))
# saveRDS(census.geo.body, "Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/working objects/shape.focusarea.rds")
census.geo.body <- readRDS("Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/working objects/shape.focusarea.rds")


```

```{r Voronoi prep, eval=F}
# adding to the coverage areas the results of the c vector
coverage.areas.comp <- coverage.areas %>% 
  left_join(c.vec, by = "antenna.ID") %>% 
  mutate(phones.sum = case_when(is.na(phones.sum) ~ 0,
                                TRUE ~ phones.sum))

# select only the internal id and the mobile phone population per tile
census.de.100m.tile.pop <- census.de.100m.tile %>% 
  dplyr::select(internal.id, pop)

```

```{r voronoi tower prep, eval=F}
# aggregating the antennas to towers (and corresponding values) and identifying problematic tower locations (outside of focus area)
tower.aggr <- coverage.areas.comp %>% 
  st_drop_geometry() %>% 
  group_by(tower.ID) %>% 
  mutate(phones.sum = sum(phones.sum)) %>% 
  distinct(tower.ID, .keep_all = T) %>% 
  ungroup() %>% 
  dplyr::select(tower.ID, X.tow, Y.tow, phones.sum) %>%
  st_as_sf(coords = c("X.tow", "Y.tow")) %>% 
  st_sf(crs = 3035) %>%  
  mutate(within.fa = lengths(st_within(., census.geo.body))) # find seeds outside the focus area

# saving towerIDs of problematic towers
towers.outside.ID <- tower.aggr %>% 
  filter(within.fa == 0) %>% 
  dplyr::select(tower.ID) %>% 
  deframe()

# filtering unproblematic towers
towers.inside <- tower.aggr %>% 
  filter(within.fa == 1)

# Finding respective nearest point on focus area border for every problematic tower location and binding the unproblematic towers back
# this object contains the final seed locations in the geometry
towers.final.data <- tower.aggr %>% 
  filter(within.fa == 0) %>%
  st_nearest_points(., census.geo.body) %>% 
  st_cast("POINT") %>% 
  .[seq(2, length(.), 2)] %>% # 
  st_as_sf() %>% 
  mutate(tower.ID = towers.outside.ID) %>% 
  mutate(geometry = x) %>% 
  st_sf(sf_column_name = "geometry") %>% 
  dplyr::select(-x) %>% 
  bind_rows(towers.inside)

sum(unlist(st_intersects(towers.final.data, census.geo.body))) == length(tower.aggr$tower.ID) # check if all are within now

# Using the seed object to calculate the Voronoi regions
tower.seed.voronoi <- towers.final.data %>%  
  st_geometry() %>% 
  st_union() %>% 
  st_voronoi() %>% 
  st_collection_extract(type = "POLYGON") %>% 
  st_sf(crs = 3035) %>% 
  st_join(tower.aggr) %>%  # rejoin with seed object to retain seed id
  st_intersection(census.geo.body) %>% 
  mutate(category.log = log10(phones.sum / as.numeric(st_area(geometry)))) %>% 
  mutate(category = cut(category.log, c(-6, -4, -2, -1), right = F, ordered_result = T))
```

And now the same preparation for the coverage area centroids (here also just "antennas" even though this is imprecise).

```{r voronoi antenna and tower prep, fig.cap="The tower centroids were used as seeds for the creation of the Voronoi regions.", eval=F}
### antennas
# aggregating the antennas to towers (and corresponding values) and identifying problematic tower locations (outside of focus area)
antenna.aggr <- coverage.areas.comp %>% 
  st_drop_geometry() %>% 
  dplyr::select(antenna.ID, geometry = antenna.centroid, phones.sum) %>%
  st_sf() %>% 
  st_sf(crs = 3035) %>%  
  mutate(within.fa = lengths(st_within(., census.geo.body))) # find seeds outside the focus area

# saving towerIDs of problematic antennas
antennas.outside.ID <- antenna.aggr %>% 
  filter(within.fa == 0) %>% 
  dplyr::select(antenna.ID) %>% 
  deframe()

# filtering unproblematic antennas
antennas.inside <- antenna.aggr %>% 
  filter(within.fa == 1)

# Finding respective nearest point on focus area border for every problematic antenna location and binding the unproblematic towers back
# this object contains the final seed locations in the geometry
antennas.final.data <- antenna.aggr %>% 
  filter(within.fa == 0) %>%
  st_nearest_points(., census.geo.body) %>% 
  st_cast("POINT") %>% 
  .[seq(2, length(.), 2)] %>% # 
  st_as_sf() %>% 
  mutate(antenna.ID = antennas.outside.ID) %>% 
  mutate(geometry = x) %>% 
  st_sf(sf_column_name = "geometry") %>% 
  dplyr::select(-x) %>% 
  bind_rows(antennas.inside)

# sum(unlist(st_intersects(antennas.final.data, census.geo.body))) == length(antenna.aggr$antenna.ID) # check if all are within now

# Using the seed object to calculate the Voronoi regions
antenna.seed.voronoi <- antennas.final.data %>%  
  st_geometry() %>% 
  st_union() %>% 
  st_voronoi() %>% 
  st_collection_extract(type = "POLYGON") %>% 
  st_sf(crs = 3035) %>% 
  st_join(antenna.aggr) %>%  # rejoin with seed object to retain seed id
  st_intersection(census.geo.body) %>% 
  mutate(category.log = log10(phones.sum / as.numeric(st_area(geometry)))) %>% 
  mutate(category = cut(category.log, c(-6, -4, -2, -1), right = F, ordered_result = T))

# check density values
# antenna.seed.voronoi %>% 
#   ggplot() +
#   geom_density(aes(category.log))

# Visualizing the Vornoi regions with respective phone count
VT.plot <- tower.seed.voronoi %>% 
  ggplot() +
  geom_sf(aes(fill = category), size = 0.001) + 
  scale_fill_ptol(na.translate = FALSE) +
  scale_color_ptol(na.translate = FALSE) +
  guides(color = F) +
  labs(fill = "Local density\n(log10)") +
  ggtitle("", subtitle = "Tower locations as seeds") +   
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 9, hjust = 0.5))

saveRDS(VT.plot, "Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/working objects/VT.plot.rds")

VA.plot <- antenna.seed.voronoi %>% 
  ggplot() +
  geom_sf(aes(fill = category), size = 0.001) + 
  scale_fill_ptol(na.translate = FALSE) +
  scale_color_ptol(na.translate = FALSE) +
  guides(color = F) +
  labs(fill = "Local density\n(log10)") +
  ggtitle("", subtitle = "Coverage area centroid locations as seeds") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1),
        plot.title = element_text(size = 10, face = "bold", hjust = 0.5),
        plot.subtitle = element_text(size = 9, hjust = 0.5))

saveRDS(VA.plot, "Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/working objects/VA.plot.rds")

plot_grid(VT.plot, VA.plot, labels = "Fig. 10: Voronoi tower estimation",  hjust = -0.1, label_size = 14)
```

We have now created the Voronoi regions for our focus area with tower locations as seeds as well as with antenna respective coverage area centroid locations as seeds. The number of seeds and Voronoi regions align with each other.

The spatial density per tile can be expressed as the ratio of the mobile phones in that region and the area of the Voronoi region:

$d_{k} = \frac{\sum c_{i}} {A_{k}}$

This region specific estimand corresponds to the tile estimate for a tile that is contained in this region. On the tile level there can be two cases that we need to differentiate:

-   Tiles that are **completely contained** within one Voronoi region 

-   Tiles that **intersect** with two or more Vornoi regions

For the first case, the tile specific estimate corresponds to $d_{k}$. For the second case we will calculate a weighted mean of the relevant Voronoi region estimates by using the corresponding tile area within each Voronoi region as weights. 

```{r Voronoi tower multiples prep, fig.cap="This is an example of a tile that intersects with multiple Voronoi regions. The labels and number correspond to the Voronoi region and the given area that are used as weights for the weighted mean estimator.", eval=F}
# spatial densities per Voronoi region
tower.voronoi.est <- tower.seed.voronoi %>% 
  mutate(area = as.numeric(st_area(.$geometry))) %>% 
  mutate(voronoi.est = phones.sum / area) # the sum of the product of this and the area is the total pop


# Joining the regions with the tile specific data
tower.seed.voronoi.join <- tower.voronoi.est %>% 
  st_join(census.de.100m.tile.pop) %>% # & re-connect the data items
  st_set_agr("aggregate") %>% # clean up
  group_by(internal.id) %>% 
  mutate(count = n()) %>% 
  ungroup()

# identifiying tiles intersecting with multiple Voronoi regions
tower.multiple <- tower.seed.voronoi.join %>%
  st_drop_geometry() %>% 
  filter(count > 1) %>% 
  distinct(internal.id) %>% 
  deframe()

# calculate area within competing voronoi regions of "multiple" tiles
tower.intersect.tiles <- census.de.100m.tile.pop %>% 
  filter(internal.id %in% tower.multiple) %>%
  st_intersection(tower.voronoi.est) %>% 
  st_collection_extract(type = "POLYGON") %>% # select the polygons
  mutate(amount.tiles = as.numeric(st_area(.$geometry)) / 10000) # checked if it adds up to 1


example.helper <- tower.intersect.tiles %>%
  filter(internal.id == "342403") %>% 
  select(tower.ID) %>% 
  deframe()
```

And now the same again for the antennas.

```{r Voronoi antenna multiples, fig.cap="This is an example of a tile that intersects with multiple Voronoi regions. The labels and number correspond to the Voronoi region and the given area that are used as weights for the weighted mean estimator.", eval=F}
# spatial densities per Voronoi region
antenna.voronoi.est <- antenna.seed.voronoi %>% 
  mutate(area = as.numeric(st_area(.$geometry))) %>% 
  mutate(voronoi.est = phones.sum / area) # the sum of the product of this and the area is the total pop


# Joining the regions with the tile specific data
antenna.seed.voronoi.join <- antenna.voronoi.est %>% 
  st_join(census.de.100m.tile.pop) %>% # & re-connect the data items
  st_set_agr("aggregate") %>% # clean up
  group_by(internal.id) %>% 
  mutate(count = n()) %>% 
  ungroup()

# identifiying tiles intersecting with multiple Voronoi regions
antenna.multiple <- antenna.seed.voronoi.join %>%
  st_drop_geometry() %>% 
  filter(count > 1) %>% 
  distinct(internal.id) %>% 
  deframe()

# calculate area within competing voronoi regions of "multiple" tiles
antenna.intersect.tiles <- census.de.100m.tile.pop %>% 
  filter(internal.id %in% antenna.multiple) %>%
  st_intersection(antenna.voronoi.est) %>% 
  st_collection_extract(type = "POLYGON") %>% # select the polygons
  mutate(amount.tiles = as.numeric(st_area(.$geometry)) / 10000) # checked if it adds up to 1

```

We can now calculate the estimated spatial density per tile.

```{r VT and VA estimate, eval=F}
#### Towers

# final datatset to calculate spatial density
tower.voronoi.final <- tower.intersect.tiles %>% 
  st_drop_geometry() %>% 
  select(internal.id, tower.ID, amount.tiles) %>% 
  right_join(tower.seed.voronoi.join, by = c("internal.id", "tower.ID")) %>% 
  mutate(amount.tiles = case_when(is.na(amount.tiles) ~ 1,
                                  TRUE ~ amount.tiles)) %>% 
  group_by(internal.id) %>% 
  mutate(voronoi.est.tower = weighted.mean(x = voronoi.est, w = amount.tiles) * 10000) %>% 
  distinct(internal.id, voronoi.est.tower) # should result in same length as the raw tiles object and the sum of the vornoir est corrected should resemble the sum of the c.vec

#### antennas

# final datatset to calculate spatial density
antenna.voronoi.final <- antenna.intersect.tiles %>% 
  st_drop_geometry() %>% 
  select(internal.id, antenna.ID, amount.tiles) %>% 
  right_join(antenna.seed.voronoi.join, by = c("internal.id", "antenna.ID")) %>% 
  mutate(amount.tiles = case_when(is.na(amount.tiles) ~ 1,
                                  TRUE ~ amount.tiles)) %>% 
  group_by(internal.id) %>% 
  mutate(voronoi.est.corrected = weighted.mean(x = voronoi.est, w = amount.tiles) * 10000) %>% 
  distinct(internal.id, voronoi.est.corrected, pop) # should result in same length as the raw tiles object and the sum of the vornoir est corrected should resemble the sum of the c.vec


voronoi.final.both <- antenna.voronoi.final %>% 
  mutate(voronoi.est.antenna = voronoi.est.corrected) %>% 
  select(-voronoi.est.corrected) %>% 
  left_join(tower.voronoi.final, by = "internal.id")

saveRDS(voronoi.final.both, "Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/Estimates/Voronoi both/Voronoi.estimates.rds")
              

```

In this chunk of code, we will then create the plots.

```{r 7gganimate,eval = F}

## data prep and plot: Voronoi Tower
estimations.Voronoi.both <- readRDS("Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/Estimates/Voronoi both/Voronoi.estimates.rds") # read in the data

estimations.VT <- estimations.Voronoi.both %>% 
  ungroup() %>% 
  select(internal.id, voronoi.est.tower) %>% 
  mutate(pop.cat = case_when(voronoi.est.tower <= 1 ~ "Less or equal 1",
                             voronoi.est.tower > 1 & voronoi.est.tower <= 10 ~ "Between 2 and 10",
                             voronoi.est.tower > 10 & voronoi.est.tower <= 30 ~ "Between 11 and 30",
                             voronoi.est.tower > 30 ~"More than 30")) %>% 
  mutate(mobile = factor(pop.cat, levels = c("Less or equal 1", "Between 2 and 10", "Between 11 and 30", "More than 30"))) %>%
  arrange(internal.id) %>% 
  mutate(geometry = census.geo$geometry) %>% 
  select(internal.id, mobile, geometry) %>% 
  st_as_sf() %>% 
  group_by(mobile) %>% 
  summarise(geometry = st_union(geometry))
saveRDS(estimations.VT, "Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/Estimates/Plot.files/VT.estimate_data.rds")

VT.plot <- estimations.VT %>% 
  ggplot() +
  geom_sf(aes(fill = factor(mobile), color = factor(mobile))) +
  labs(x = "", y = "", title = "", subtitle = "Voronoi tower estimate", size = 20) +
  scale_fill_ptol(name = "Mobile phones") +
  scale_color_ptol(guide = F)
saveRDS(VT.plot, "Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/Estimates/Plot.files/VT.estimate_map.rds")


## data prep and plot: Voronoi antenna
estimations.Voronoi.both <- readRDS("Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/Estimates/Voronoi both/Voronoi.estimate.rds")

estimations.VA <- estimations.Voronoi.both %>% 
  ungroup() %>% 
  select(internal.id, voronoi.est.antenna) %>% 
  mutate(pop.cat = case_when(voronoi.est.antenna <= 1 ~ "Less or equal 1",
                             voronoi.est.antenna > 1 & voronoi.est.antenna <= 10 ~ "Between 2 and 10",
                             voronoi.est.antenna > 10 & voronoi.est.antenna <= 30 ~ "Between 11 and 30",
                             voronoi.est.antenna > 30 ~ "More than 30")) %>% 
  mutate(mobile = factor(pop.cat, levels = c("Less or equal 1", "Between 2 and 10", "Between 11 and 30", "More than 30"))) %>%
  arrange(internal.id) %>% 
  mutate(geometry = census.geo$geometry) %>% 
  select(internal.id, mobile, geometry) %>% 
  st_as_sf() %>% 
  group_by(mobile) %>% 
  summarise(geometry = st_union(geometry))
saveRDS(estimations.VA, "Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/Estimates/Plot.files/VA.estimate_data.rds")

VA.plot <- estimations.VA %>% 
  ggplot() +
  geom_sf(aes(fill = factor(mobile), color = factor(mobile))) +
  labs(x = "", y = "", title = "", subtitle = "Voronoi antenna estimate", size = 20) +
  scale_fill_ptol(name = "Mobile phones") +
  scale_color_ptol(guide = F)
saveRDS(VA.plot, "Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/Estimates/Plot.files/VA.estimate_map.rds")

```

```{r voronoi antenna and tower prep plot, fig.cap="The tower centroids were used as seeds for the creation of the Voronoi regions. The color scale is defined by the log with base 10 local density i.e. the absolute number of mobile phones within a Voronoi region divided by absolute area of the respecitve Voronoi region.",fig.show='hold'}

# Plot read in
VT.plot <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/working objects/VT.plot.rds")
VA.plot <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/working objects/VA.plot.rds")

VT.plot
VA.plot

# plot_grid(VT.plot, VA.plot, labels = "Fig. 10: Voronoi tower estimation",  hjust = -0.1, label_size = 14)

```

We have now created the Voronoi regions for our focus area with tower locations as seeds as well as with antenna respective coverage area centroid locations as seeds. The number of seeds and Voronoi regions align with each other.

The spatial density per tile can be expressed as the ratio of the mobile phones in that region and the area of the Voronoi region:

$d_{k} = \frac{\sum c_{i}} {A_{k}}$

This region specific estimate corresponds to the tile estimate for a tile that is contained in this region. On the tile level there can be two cases that we need to differentiate:

-   Tiles that are completely contained within one Voronoi region 

-   Tiles that intersect with two or more Vornoi regions

For the first case, the tile specific estimate corresponds to $d_{k}$. For the second case we will calculate a weighted mean of the relevant Voronoi region estimates by using the corresponding tile area within each Voronoi region as weights. 

``` {r bird plot, eval=F}

# bird example plot
VT.bird.example.plot <- tower.intersect.tiles %>%
  filter(internal.id == "342403") %>%
  ggplot() +
  geom_sf(color = "black") +
  geom_sf(data = (census.de.100m.tile %>% filter(internal.id == "342403")), color = "red", fill = NA) +
  geom_sf(data = (tower.seed.voronoi %>% filter(tower.ID %in% example.helper)), fill = NA) +
  ggtitle("", subtitle = "Birds view of a tile that intersects with 
multiple regions")

saveRDS(VT.bird.example.plot, "Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/working objects/VT.bird.example.plot.rds")

VT.example.plot <- tower.intersect.tiles %>% 
  filter(internal.id == "342403") %>% 
  ggplot() +
  geom_sf() + # add description and scales
  geom_sf_label(aes(label = tower.ID)) +
  geom_sf_text(aes(label = paste0("(", round(amount.tiles, 2), ")")), nudge_y = -7) +
  geom_sf(data = (census.de.100m.tile.pop %>% 
  filter(internal.id == "342403")), color = "red", fill = NA) +
  ggtitle("", subtitle = "Relative area per Voronoi region")

saveRDS(VT.example.plot, "Vysoká škola ekonomická v Praze/Tony Wei Tse Hung - YAY/working objects/VT.example.plot.rds")

plot_grid(VT.bird.example.plot, VT.example.plot, 
          labels = "Fig. 11: Example multiple tile",  hjust = -0.1, label_size = 14)

```

```{r Voronoi tower multiples, fig.cap="This is an example of a tile that intersects with multiple Voronoi regions. The labels and number correspond to the Voronoi region and the given area that are used as weights for the weighted mean estimator.",fig.show='hold'}

VT.bird.example.plot <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/working objects/VT.bird.example.plot.rds") + theme(axis.text.x=element_text(angle=90, hjust=1))
VT.example.plot <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/working objects/VT.example.plot.rds") + theme(axis.text.x=element_text(angle=90, hjust=1))

VT.bird.example.plot
VT.example.plot

# plot_grid(VT.bird.example.plot, VT.example.plot, 
#           labels = "Fig. 11: Example multiple tile",  hjust = -0.1, label_size = 14)
```

Here we visualize the second case with an example tile and example Voronoi regions from the Voronoi with tower locations as seeds. In the left plot we can see a small tile that is indicated in red. This tile is at the intersection of 4 Voronoi regions. The corresponding relative tile area intersecting with each Voronoi region can be seen in the right plot. The labels correspond to the Voronoi region ID and the below in brackets indicate the proportional weight that should be considered in calculting the tile specific weighted mean estimate. To compare this to the first case - where a tile is fully contained only in one region - the weight for the particular Voronoi region is 1. The same methodogolgy is pursued for the Voronoi antenna case.


## MLE Poisson

In order to iteratively calculate the MLE Poisson, we would first build a modified version of the model matrix used in the device-to-cell generation process. The modification of the model matrix is just a slightly perturbed version of the original matrix. We then used the perturbed matrix to estimate the present population in each cell. 

MLE Poisson: $\hat{u}^{m+1}_{j} = \hat{u}^{m}_{j}\sum_{i = 1}^{I} c_{i} \frac{p_{ij}}{\sum_{k = 1}^{J} p_{ik} \hat{u}^{m}_{k}}$

With the above formula, we will input different prior vectors (initial guess) to test the behavior of this estimation. We will use a non-informative prior (also called equal matrix in our book) which means that we will compute directly from the model matrix, without inserting an initial guess. Another prior vector we will use is called the informative prior (also called true or oracle matrix in our notebook) which means that we use the information given in the pertubed matrix 

This estimation works by iteratively estimating the previous estimation and slowly improves over many iterations. E.g. a block with a population of 10 and the first iteration will result in 0.3 and then the 300th iteration it will estimate a population of 4. The estimations are considered stable when the next iteration is no longer changing from the previous estimation. j