# Appendix A: A deep dive into the convergence behaviour of MLE estimates


```{r packages, message=FALSE, echo = F}
library(tidyverse)
library(sf)
library(raster)
library(Matrix)
library(knitr)
library(kableExtra)
library(ggthemes)
library(cowplot)
library(transformr)
library(gganimate)
set.seed(762)
```

In the course of our work we thought it was particularly interesting to further focus on the actual convergence behavior of the MLE-based estimators. By this, we want to see how the estimators evolve after each iteration.

We will visualize the convergence behavior/the evolution of the estimates up to 1000 iterations. This is executed based on a stratified sample (strata: three level `pop.area.kind` variable), n = 150). We measure this convergence behavior through the use of MAD and AAD which was explained in chapter 5. We calculate the AAD and MAD for each iteration and see how estimates evolve, as the iterations increase.

-   Average Absolute Discrepancy (AAD): $D_{avg}(u,v) = \frac{1}{U}\sum_{j = 1}^{J} |u_{j} - v_{j}|$

-   Maximum Absolute Discrepancy (MAD): $D_{max}(\textbf{u,v}) \overset{\underset{\mathrm{def}}{}}{=} \sum_{\mathit{j = 1}}^{\mathit{J}}\left | u_{j} - v_{j} \right |$

## Discrepancy measures per iteration

``` {r 5BAADMAD, eval = F}

# In this section we will be calculating the AAD and MAD values for the true population matrix

# First get the true population vector from another dataset
voro <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/Estimates/Voronoi both/Voronoi.estimates.rds") # getting the true pop vector from another object
pop = voro$pop
voro = voro %>% ungroup()

# Calculating Average Absolute Discrepancy
# Because the data is in three parts, thus we will calculate the AAD separately. The first set has iterations between 0 and 400. The second set has iterations between 401 and 801. The third set has iterations between 802 and 1002. 
# The code/procedure is rather similar to the statistical evaluation presented in chapter 5, except that over here we are looking at all of the iterations instead of the last iteration.
# Note: the graphing part might take a bit longer to render.

# Part one
equal1 <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/Estimates/P.equal/u.est.non.inf.P.equal.part1_400_new.rd s" ) #read in the data
equal1 = cbind(equal1, pop) # combining the data frame with the population vector
bigu = sum(equal1$pop) # useful to have this calculated and stored as a constant already 
aad1 = equal1 %>%  
  transmute_at(vars(contains("u")),funs("diff" = abs(pop - .))) %>% # transforming the columns
  summarise_all(funs("i" = sum(.)/bigu)) # summing up all rows for each column
colnames(aad1) = 0:400 # change the column names

# Part two
equal2 <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/Estimates/P.equal/u.est.non.inf.P.equal.part401_800_new.rds") # read in the data
equal2 = cbind(equal2,pop) # combining the data frame with the population vector
aad2 = equal2 %>%  # mutating 
  transmute_at(vars(contains("u")),funs("diff" = abs(pop - .))) %>% 
  summarise_all(funs("i" = sum(.)/bigu))
colnames(aad2) = 401:801 # change the column names

# Part three
equal3 <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/Estimates/P.equal/u.est.non.inf.P.equal.part801_1000_new.rds") # read in the data 
equal3 = cbind(equal3,pop) # combining the data frame with the population vector
aad3 = equal3 %>%  
  transmute_at(vars(contains("u")),funs("diff" = abs(pop - .))) %>% # transforming the columns
  summarise_all(funs("i" = sum(.)/bigu)) # summing up all rows for each column
colnames(aad3) = 802:1002 # change the column names

aad = cbind(aad1,aad2,aad3) # now we merge all three parts together in one
aadt = aad %>%
  pivot_longer(-c(), names_to = "iter", values_to = "aad") # transform the data frame into the long format

aadt$iter = as.numeric(aadt$iter) # cast the `iter` variable as a numeric variable

# Plot using ggplot
equal.aad.plot = aadt %>% 
  ggplot(aes(x = iter, y = aad)) + 
  geom_line() + labs(x = "Number of Iteration", y = "Average Absolute Discrepancy", title = "Equal Probability Matrix ~ AAD") + theme(axis.text.x = element_text(angle = 90),legend.position = "bottom") +
  scale_x_continuous(breaks=seq(0, 1000, 50))

saveRDS(equal.aad.plot, file = "equal.aad.plot.rds") # save the plot as rds

# Calculating Maximum Absolute Discrepancy

# Part One
mad1 = equal1 %>%  
  transmute_at(vars(contains("u")),funs("diff" = abs(pop - .))) %>% # transforming the columns
  summarise_all(funs("i" = sum(.))) # summing up all rows for each column
colnames(mad1) = 0:400 # change the column names

# Part Two
mad2 = equal2 %>%  # mutating 
  transmute_at(vars(contains("u")),funs("diff" = abs(pop - .))) %>% # transforming the columns
  summarise_all(funs("i" = sum(.))) # summing up all rows for each column
colnames(mad2) = 401:801 # change the column names

# Part Three
mad3 = equal3 %>%  # mutating 
  transmute_at(vars(contains("u")),funs("diff" = abs(pop - .))) %>% # transforming the columns
  summarise_all(funs("i" = sum(.))) # summing up all rows for each column
colnames(mad3) = 802:1002 # change the column names

mad = cbind(mad1,mad2,mad3) # now we merge all three parts together in one
madt = mad %>%
  pivot_longer(-c(),names_to = "iter", values_to = "mad") # transform the data frame into the long format

madt$iter = as.numeric(madt$iter) # cast the `iter` variable as a numeric variable

# Plot using ggplot
equal.mad.plot = madt %>% 
  ggplot(aes(x = iter, y = mad)) + 
  geom_line() + labs(x = "Number of Iteration", y = "Maximum Absolute Discrepancy", title = "Equal Probability Matrix ~ MAD") + theme(axis.text.x = element_text(angle = 90),legend.position = "bottom") +
  scale_x_continuous(breaks=seq(0, 1000, 50))

saveRDS(equal.mad.plot, file = "equal.mad.plot.rds") # save the plot as rds

```

```{r equal aad/mad, echo = F}

aad.plot <- readRDS(url("https://github.com/R-ramljak/MNO_Eurostat/raw/master/Plots/AADMAD/equal.aad.plot.rds", method = "libcurl")) + ggtitle("", subtitle = "AAD")

mad.plot <- readRDS(url("https://github.com/R-ramljak/MNO_Eurostat/raw/master/Plots/AADMAD/equal.mad.plot.rds", method = "libcurl")) + ggtitle("", subtitle = "MAD")

plot_grid(aad.plot, mad.plot,
          labels = "Fig. 17 Discrepancy Over Time, Equal Pop. Matrix", 
          label_size = 12, align = "hv", hjust = -0.1)
```
For the equal population matrix, the discrepancy values stabilizes as the number of iterations increases. Specifically, starting at the 700th iteration, the discrepancy values are largely the same. 

``` {r, eval = F}

# In this section we will be calculating the AAD and MAD values for the true population matrix.
# This part is the largely the same procedure as the above section when calculating the equal population matrix. 

# NOte: due to memory shortage, I have actively remove unused objects in order to conserve as much memory as possible. I also used as much of the garbage collection function, gc(), as possible. 

rm(list = ls()) # remove unused objects

voro <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/Estimates/Voronoi both/Voronoi.estimates.rds") # getting the true pop vector from another object
pop = voro$pop
rm(voro)

# Calculating Average Absolute Discrepancy

# Part One
true1 <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/Estimates/P.oracle/u.est.non.inf.P.oracle.part1_400_new.rds") # read in the data
true1 = cbind(true1,pop) # combining the data frame with the population vector
bigu = sum(true1$pop) # useful to have this calculated and stored as a constant already 
aad1 = true1 %>%  
  transmute_at(vars(contains("u")),funs("diff" = abs(pop - .))) %>% # transforming the columns
  summarise_all(funs("i" = sum(.)/bigu)) # summing up all rows for each column
colnames(aad1) = 0:400 # change column names
rm(true1) # remove unnecessary object

# Part Two
true2 <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/Estimates/P.oracle/u.est.non.inf.P.oracle.part401_800_new.rds") # read in the data
true2 = cbind(true2,pop) # combining the data frame with the population vector
aad2 = true2 %>%  
  transmute_at(vars(contains("u")),funs("diff" = abs(pop - .))) %>% # transforming the columns
  summarise_all(funs("i" = sum(.)/bigu)) # summing up all rows for each column
colnames(aad2) = 401:801# change column names
rm(true2)# remove unnecessary object

# Part Three
true3 <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/Estimates/P.oracle/u.est.non.inf.P.oracle.part801_1000_new.rds") # read in the data
true3 = cbind(true3,pop) # combining the data frame with the population vector
aad3 = true3 %>%  
  transmute_at(vars(contains("u")),funs("diff" = abs(pop - .))) %>% # transforming the columns
  summarise_all(funs("i" = sum(.)/bigu)) # summing up all rows for each column
colnames(aad3) = 802:1002 # change column names
rm(true3)# remove unnecessary object

aad = cbind(aad1,aad2,aad3) # combining all three parts together in one data frame
aadt = aad %>%
  pivot_longer(-c(), names_to = "iter", values_to = "aad") # Transform the data frame into long form

aadt$iter = as.numeric(aadt$iter) # cast the `iter` variable as a numeric variable

# Plot using ggplot
true.aad.plot = aadt %>% 
  ggplot(aes(x = iter, y = aad)) + 
  geom_line() + labs(x = "Number of Iteration", y = "Average Absolute Discrepancy", title = "True Probability Matrix ~ AAD") + theme(axis.text.x = element_text(angle = 90),legend.position = "bottom") +
  scale_x_continuous(breaks=seq(0, 1000, 50))

saveRDS(true.aad.plot, file = "true.aad.plot.rds") # save the plot as a rds object

# Calculating Maximum Absolute Discrepancy

rm(list = ls()) # remove unnecessary objects

voro <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/Estimates/Voronoi both/Voronoi.estimates.rds") # getting the true pop vector from another object
pop = voro$pop
rm(voro)

# Part One
true1 <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/Estimates/P.oracle/u.est.non.inf.P.oracle.part1_400_new.rds") # read in the data
true1 = cbind(true1,pop) # combining the data frame with the population vector
mad1 = true1 %>% 
  transmute_at(vars(contains("u")),funs("diff" = abs(pop - .))) %>% # transforming the columns
  summarise_all(funs("i" = sum(.))) # summing up all rows for each column
colnames(mad1) = 0:400 # change column names
rm(true1) # remove unused object

# Part Two
true2 <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/Estimates/P.oracle/u.est.non.inf.P.oracle.part401_800_new.rds") # read in the data
true2 = cbind(true2,pop) # combining the data frame with the population vector
mad2 = true2 %>% 
  transmute_at(vars(contains("u")),funs("diff" = abs(pop - .))) %>% # transforming the columns
  summarise_all(funs("i" = sum(.))) # summing up all rows for each column
colnames(mad2) = 401:801 # change column names
rm(true2) # remove unused object

# Part Three
true3 <- readRDS("~/OneDrive - Vysoká škola ekonomická v Praze/YAY/Estimates/P.oracle/u.est.non.inf.P.oracle.part801_1000_new.rds") # read in the data
true3 = cbind(true3,pop) # combining the data frame with the population vector
mad3 = true3 %>%  
  transmute_at(vars(contains("u")),funs("diff" = abs(pop - .))) %>% # transforming the columns
  summarise_all(funs("i" = sum(.))) # summing up all rows for each column
colnames(mad3) = 802:1002 # change column names
rm(true3) # remove unused object

mad = cbind(mad1,mad2,mad3)  # combining all three parts together in one data frame
madt = mad %>%
  pivot_longer(-c(),names_to = "iter", values_to = "mad") # Transform the data frame into long form

madt$iter = as.numeric(madt$iter) # cast the `iter` variable into numeric

# plot using ggplot
true.mad.plot = madt %>% 
  ggplot(aes(x = iter, y = mad)) + 
  geom_line() + labs(x = "Number of Iteration", y = "Maximum Absolute Discrepancy", title = "True Probability Matrix ~ MAD") + theme(axis.text.x = element_text(angle = 90),legend.position = "bottom") +
  scale_x_continuous(breaks=seq(0, 1000, 50))

saveRDS(true.mad.plot, file = "true.mad.plot.rds") # save the plot as a rds object

rm(list = ls()) # remove 

```


```{r true aad/mad, echo = F}
aad.plot <- readRDS(url("https://github.com/R-ramljak/MNO_Eurostat/raw/master/Plots/AADMAD/true.aad.plot.rds", method = "libcurl")) + ggtitle("", subtitle = "AAD")

mad.plot <- readRDS(url("https://github.com/R-ramljak/MNO_Eurostat/raw/master/Plots/AADMAD/true.mad.plot.rds", method = "libcurl")) + ggtitle("", subtitle = "MAD")

plot_grid(aad.plot, mad.plot,
          labels = "Fig. 18 Discrepancy Over Time, True Pop. Matrix", 
          label_size = 12, align = "hv", hjust = -0.1)
```
For the true population matrix, the discrepancy values stabilizes as the number of iterations increases. More specifically, starting at the 300th iteration, the discrepancy values are largely the same. 

## Convergence Behavior

In this part, we will look at the convergence behavior over time through the actual numbers of estimated population and the residual. The residual is calculated as the *true population minus the estimated population*. The residual plot only contain certain number of iterations. 

``` {r, eval = F}

esample <- readRDS("~/Desktop/Eurostat/est.equal.1000iter.sample.rds") # Load in the data
esample$internal.id = as.factor(esample$internal.id) # recast the `internal id` as a factor
esample = esample %>% select(-c(u.true.y)) # take out u.true.y
esample = esample %>% select(-c(pop.area.kind.helper,u.true)) # take out other columns
length(colnames(esample)[406:1007]) # checking the length of the column names

# Generate the column names through the use of for loops
# It's important to put 0 in front or else ggplot will then graph 1000 in front of 100, for example.
iters = c() 
for (x in 0:9) {
  iters = c(iters, paste("000",x,sep = ""))
}
for (x in 10:99) {
  iters = c(iters, paste("00",x,sep = ""))
}
for (x in 100:400) {
  iters = c(iters, paste("0",x,sep = ""))
}
colnames(esample)[2:402] = iters
iters2 = c()
for (x in 401:999) {
  iters2 = c(iters2, paste("0",x,sep = ""))
}
for (x in 1000:1002){
  iters2 = c(iters2, paste("",x,sep = ""))
}
colnames(esample)[406:1007] = iters2

equal = esample %>% # transform the data frame into long form
  select(-c(pop)) %>% 
  pivot_longer(-c(internal.id, pop.area.kind,u.true.x), names_to = "iter", values_to = "sim")

equal$iter = as.numeric(equal$iter) # cast the `iter` variable as a numeric variable

# Plot of Estimated Population over Iterations
equal.plot = equal %>% 
  ggplot(aes(x = iter, y = sim, group = internal.id)) + 
  geom_line(aes(color = pop.area.kind)) + 
  labs(x = "Number of Iteration", y = "Estimated Population", title = "Equal Probability Matrix") + 
  theme(axis.text.x = element_text(angle = 90),legend.position = "bottom") + 
  scale_color_ptol(breaks = c("Rural", "Suburban", "Urban")) + 
  facet_grid(vars(pop.area.kind),scales="free") +
  scale_x_continuous(breaks=seq(0, 1000, 50))

saveRDS(equal.plot, file = "equal.plot.rds") # save the plot as RDS

# Residual Plot

esample <- readRDS("~/Desktop/Eurostat/est.equal.1000iter.sample.rds") # read in the data
esample = esample %>% select(-c(u.true.y,pop.area.kind.helper)) # take out unused columns

# Generate column names through the use of for loops
niters = c() 
for (x in 0:9) {
  niters = c(niters, paste("u000",x,sep = ""))
}
for (x in 10:99) {
  niters = c(niters, paste("u00",x,sep = ""))
}
for (x in 100:400) {
  niters = c(niters, paste("u0",x,sep = ""))
}
colnames(esample)[2:402] = niters
niters2 = c()
for (x in 401:999) {
  niters2 = c(niters2, paste("u0",x,sep = ""))
}
for (x in 1000:1002){
  niters2 = c(niters2, paste("u",x,sep = ""))
}
colnames(esample)[406:1007] = niters2

# Calculate the residuals and transform the data frame into the long format.
equalr = esample %>% 
  mutate(r0001 = u.true - u0001, r0002 = u.true - u0002, r0005 = u.true - u0005,
         r0010 = u.true - u0010, r0020 = u.true - u0020, r0050 = u.true - u0050,
         r0100 = u.true - u0100, r0200 = u.true - u0200, r0300 = u.true - u0300,
         r0400 = u.true - u0400, r0500 = u.true - u0500, r0600 = u.true - u0600,
         r0700 = u.true - u0700, r0800 = u.true - u0800, r0900 = u.true - u0900,
         r1000 = u.true - u1000
         ) %>%
  select(pop.area.kind,internal.id,r0001,r0002,r0005,r0010,r0020,r0050,r0100,r0200,r0300,r0400,r0500,r0600,r0700,r0800,r0900,r1000) %>%
  pivot_longer(-c(internal.id, pop.area.kind), names_to = "iter", values_to = "resid")

# Plot using ggplot  
equal.resid.plot = equalr %>%
  ggplot(aes(x = iter, y = resid, group = internal.id)) + 
  geom_line(aes(color = pop.area.kind)) + labs(x = "Number of Iteration", y = "Residual = Actual - Predicted", title = "Equal Probability Matrix ~ Residual") + theme(axis.text.x = element_text(angle = 90),legend.position = "bottom") + scale_color_ptol(breaks = c("Rural", "Suburban", "Urban")) + 
  facet_grid(vars(pop.area.kind),scales = "free") 

saveRDS(equal.resid.plot, file = "equal.resid.plot.rds") # save the plot as ggplot. 

```


```{r distributionl eval 1, echo = F, fig.cap= "The left and right plots show a stratified sample (strata: area kind) and the tile specific estimatands' convergence behaviour. The convergence behaviour differs between area kinds. The right plot emphasizes that the true value is not necessarily reached."}

equal.plot <- readRDS(url("https://github.com/R-ramljak/MNO_Eurostat/raw/master/Plots/Estimation%20Plots/equal.plot.rds", method = "libcurl")) + ggtitle("", subtitle = "Evolution")
equal.resid.plot <- readRDS(url("https://github.com/R-ramljak/MNO_Eurostat/raw/master/Plots/Estimation%20Plots/equal.resid.plot.rds", method = "libcurl")) + ggtitle("", subtitle = "Residual")

plot_grid(equal.plot, equal.resid.plot,
          labels = "Fig. 19 Convergence behaviour (Equal P.matrix)", 
          label_size = 12, hjust = -0.1) 

```

The graphs above shows how the estimated population evolve over time (left) as well as its prediction error (right). From the graph on the left, most of the estimated population stabilizes rather quickly in rural and suburban tiles. With the urban tiles, it is harder to say exactly when the convergence will occur, especially there are a bit of variation. The residual plot (right) shows the error of prediction in comparison to the actual population in that specific tile. The rural tiles shows a trend of constant error of prediction over iterations. The suburban tiles generally shows the same trend as well, but also shows that there are tiles that overestimated. The urban tiles tend to have an underestimation.

``` {r 5A_EP, eval = F}

tsample <- readRDS("~/Desktop/Eurostat/est.true.1000iter.sample.rds") # read in data
tsample$internal.id = as.factor(tsample$internal.id) # cast the `internal id` variable as a factor
colnames(tsample)

colnames(tsample)[2:402] = iters # reuse the column names generated above
colnames(tsample)[405:1006] = iters2

tequal = tsample %>% # transform the data frame into long form
  pivot_longer(-c(internal.id, pop.area.kind,pop), names_to = "iter", values_to = "sim") 

tequal$iter = as.numeric(tequal$iter) # cast the `iter` variable as a numeric variable

# Estimated Population over Iterations plot using ggplot
true.plot = ggplot(data = tequal, aes(x = iter, y = sim, group = internal.id)) + 
  geom_line(aes(color = pop.area.kind)) + labs(x = "Number of Iteration", y = "Estimated Population",     title = "True Population Matrix") + 
  theme(axis.text.x = element_text(angle = 90),legend.position = "bottom") +
  scale_color_ptol(breaks = c("Rural", "Suburban", "Urban")) + 
  facet_grid(vars(pop.area.kind), scales = "free") +
  scale_x_continuous(breaks=seq(0, 1000, 50))

saveRDS(true.plot, file = "true.plot.rds") # save the plot as rds

# Residual Plot

colnames(tsample)[2:402] = niters # reused column names from above
colnames(tsample)[405:1006] = niters2

# Calculate the residuals and transform the data frame into the long format.
tequalr = tsample %>% 
  mutate(r001 = pop - u0001, r002 = pop - u0002, r005 = pop - u0005,
         r010 = pop - u0010, r020 = pop - u0020, r050 = pop - u0050,
         r100 = pop - u0100, r200 = pop - u0200, r300 = pop - u0300,
         r400 = pop - u0400, r500 = pop - u0500, r600 = pop - u0600,
         r700 = pop - u0700, r800 = pop - u0800, r900 = pop - u0900,
         r1000 = pop - u1000
  ) %>%
  select(pop.area.kind,internal.id,r001,r002,r005,r010,r020,r050,r100,r200,r300,r400,r500,r600,r700,r800) %>%
  pivot_longer(-c(internal.id, pop.area.kind), names_to = "iter", values_to = "resid")

# Plot using ggplot
true.resid.plot = ggplot(data = tequalr, aes(x = iter, y = resid, group = internal.id)) + 
  geom_line(aes(color = pop.area.kind)) + labs(x = "Number of Iteration", y = "Residual = Actual - Predicted", title = "True Population Matrix ~ Residual") + 
  theme(axis.text.x = element_text(angle = 90),legend.position = "bottom") + 
  scale_color_ptol(breaks = c("Rural", "Suburban", "Urban")) + 
  facet_grid(vars(pop.area.kind),scales = "free") 

saveRDS(true.resid.plot, file = "true.resid.plot.rds") # save ggplot as rds format


```

```{r distributionl eval 2, echo = F, fig.cap= "The left and right plots show a stratified sample (strata: area kind) and the tile specific estimatands' convergence behaviour (equal probability P.matrix). The convergence behaviour differs between area kinds. The right plot emphasizes that the true value is not necessarily reached."}

true.plot <- readRDS(url("https://github.com/R-ramljak/MNO_Eurostat/raw/master/Plots/Estimation%20Plots/true.plot.rds", method = "libcurl")) + ggtitle("", subtitle = "Evolution")

true.resid.plot <- readRDS(url("https://github.com/R-ramljak/MNO_Eurostat/raw/master/Plots/Estimation%20Plots/true.resid.plot.rds", method = "libcurl")) + ggtitle("", subtitle = "Residual")

plot_grid(true.plot, true.resid.plot,
          labels = "Fig. 20 Convergence behaviour (True P.matrix)", 
          label_size = 12, align = "hv", hjust = -0.1) 

```

Using the true population matrix, the estimated population stabilizes faster over time in comparison to the equal population matrix. However, there are a few tiles that do not stabilize as fast as we hoped. This is particularly evident in the urban tiles. The residual plot (right) shows the error of prediction over time by the type of tile. There seems to be a trend of increasing prediction error as the degree of urbanization increases as well. We can see that the range of error for suburban tiles range from 30 to -10 and the range of error for urban tiles are between 100 and -75. The suburban tiles seems to underestimate and the urban tiles seems to overestimate.